### 2020-05-14 pldi/microbenchmarks-table.tex
Update table and paper plots

\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} lllllll}
\toprule
Script ~&~ Structure & Input &Seq Time & \#Nodes(16, 64) &Compile Time (16, 64) & Highlights \\
\midrule
\tti{grep} ~&~ $3\times$\tsta & 1~GB & 79m35.197s & 49\qquad 193 & 0.056s\qquad 0.523s & complex NFA regex \\
\tti{sort} ~&~ $\tsta, \tpur$ & 10~GB & 21m46.807s & 77\qquad 317 & 0.090s\qquad 1.083s & \tti{sort}ing \\
\tti{top-n} ~&~ $2\times\tsta, 4\times\tpur$ & 10~GB & 78m45.872s & 96\qquad 384 & 0.145s\qquad 1.790s & double \tti{sort}, \tti{uniq} reduction \\
\tti{wf} ~&~ $3\times\tsta, 3\times\tpur$ & 10~GB & 22m30.048s & 96\qquad 384 & 0.147s\qquad 1.809s & double \tti{sort}, \tti{uniq} reduction \\
\tti{grep-light} ~&~ $3\times$\tsta & 100~GB & 1m38.212s & \todo{X}\qquad \todo{X} & 0.031s\qquad 0.163s & 	odo{light computation} \\
\tti{spell} ~&~ $4\times\tsta, 3\times\tpur$ & 1~GB & 8m1.428s & 69\qquad 261 & 0.104s\qquad 1.060s & comparisons (\tti{comm}) \\
\tti{shortest-scripts} ~&~ $5\times\tsta, 2\times\tpur$ & 85~MB & 28m25.818s & \todo{X}\qquad \todo{X} & 0.210s\qquad 2.762s & \todo{extensive file-system operation} \\
\tti{diff} ~&~ $2\times\tsta, 3\times\tpur$ & 10~GB & 25m10.092s & \todo{X}\qquad \todo{X} & 0.254s\qquad 1.958s & non-parallelizable \tti{diff}ing \\
\tti{optimized bi-grams} ~&~ $3\times$\tsta, \tpur & 10~GB & 125m10.224s & \todo{X}\qquad \todo{X} & 0.098s\qquad 1.171s & optimized version of bigrams \\
\tti{set-diff} ~&~ $5\times\tsta, 2\times\tpur$ & 10~GB & 51m22.068s & \todo{X}\qquad \todo{X} & 0.250s\qquad 3.316s & two pipelines merging to a \tti{comm} \\
\bottomrule
\end{tabular*}