### 2020-09-21 evaluation/plots/coarse-microbenchmarks-table.tex
Change some one-liner names to not confuse reviewers that they are simple commands

\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} llllllll}
\toprule
Script ~&~ Structure & Input &Seq. Time & Highlights \\
\midrule
nfa-regex ~&~ $3\times\tsta$ & 1~GB & 79m35.197s & complex NFA regex \\
sort ~&~ $\tsta, \tpur$ & 10~GB & 21m46.807s & \tti{sort}ing \\
top-n ~&~ $2\times\tsta, 4\times\tpur$ & 10~GB & 78m45.872s & double \tti{sort}, \tti{uniq} reduction \\
wf ~&~ $3\times\tsta, 3\times\tpur$ & 10~GB & 22m30.048s & double \tti{sort}, \tti{uniq} reduction \\
spell ~&~ $4\times\tsta, 3\times\tpur$ & 3~GB & 25m7.560s & comparisons (\tti{comm}) \\
difference ~&~ $2\times\tsta, 3\times\tpur$ & 10~GB & 25m49.097s & non-parallelizable \tti{diff}ing \\
bi-grams ~&~ $3\times\tsta, 3\times\tpur$ & 3~GB & 38m9.922s & stream shifting and merging \\
set-difference ~&~ $5\times\tsta, 2\times\tpur$ & 10~GB & 51m32.313s & two pipelines merging to a \tti{comm} \\
shortest-scripts ~&~ $5\times\tsta, 2\times\tpur$ & 85~MB & 28m45.900s & long \tsta pipeline ending with \tpur \\
\bottomrule
\end{tabular*}