### 2020-11-12 evaluation/plots/coarse-microbenchmarks-table.tex
Remove mkfifo from table

\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} l}
\toprule
Name ~&~ Script \\
\midrule
nfa-regex ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | tr A-Z a-z | grep '\(.\).*\1\(.\).*\2\(.\).*\3\(.\).*\4'! \\
sort ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | tr A-Z a-z | sort! \\
top-n ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | tr -cs A-Za-z '\n' | tr A-Z a-z | sort | uniq -c | sort -rn | sed ${N}q! \\
wf ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | tr -cs A-Za-z'\n' | tr A-Z a-z | sort | uniq -c | sort -rn | sed ${1}q! \\
spell ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | col -bx |tr -cs A-Za-z '\n' | tr A-Z a-z |tr -d '[:punct:]' |sort |uniq |comm -23 - $dict! \\
difference ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | sort | tr [:lower:] [:upper:] > s1 & cat $IN | sort | tr [:upper:] [:lower:] > s2 & diff -B s1 s2! \\
bi-grams ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | tr -cs A-Za-z '\n' | tr A-Z a-z | bigrams_aux | sort | uniq! \\
set-difference ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | cut -d ' ' -f 1 | sort > s1 & cat $IN | cut -d ' ' -f 1 | sort > s2 & comm -23 s1 s2! \\
shortest-scripts ~&~ \lstinline[columns=fixed,basicstyle=\footnotesize\ttfamily]!cat $IN | xargs file | grep "shell script" | cut -d: -f1 | xargs -L 1 wc -l | grep -v '^0$' | sort -n | head -15! \\
\bottomrule
\end{tabular*}