### 2021-03-18 evaluation/results_archive/plots/microbenchmarks-tablesmall.tex

\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} llllllll}
\toprule
Script ~&~ Structure & Input &Seq. Time & \multicolumn{2}{l}{\#Nodes(16, 64)} &\multicolumn{2}{l}{Compile Time (16, 64)} & Highlights \\
\midrule
nfa-regex ~&~ $3\times\tsta$ & 1~GB & 7m59.843s & \todo{X} & \todo{X} & 0.000s & 0.000s & complex NFA regex \\
sort ~&~ $\tsta, \tpur$ & 10~GB & 2m6.756s & \todo{X} & \todo{X} & 0.000s & 0.000s & \tti{sort}ing \\
top-n ~&~ $2\times\tsta, 4\times\tpur$ & 10~GB & 7m6.963s & \todo{X} & \todo{X} & 0.000s & 0.000s & double \tti{sort}, \tti{uniq} reduction \\
wf ~&~ $3\times\tsta, 3\times\tpur$ & 10~GB & 2m10.746s & \todo{X} & \todo{X} & 0.000s & 0.000s & double \tti{sort}, \tti{uniq} reduction \\
spell ~&~ $4\times\tsta, 3\times\tpur$ & 3~GB & 46.520s & 195 & \todo{X} & 0.324s & 0.000s & comparisons (\tti{comm}) \\
difference ~&~ $2\times\tsta, 3\times\tpur$ & 10~GB & 2m35.132s & \todo{X} & \todo{X} & 0.000s & 0.000s & non-parallelizable \tti{diff}ing \\
bi-grams ~&~ $3\times\tsta, 3\times\tpur$ & 3~GB & 1m11.240s & 284 & \todo{X} & 0.301s & 0.000s & stream shifting and merging \\
set-difference ~&~ $5\times\tsta, 2\times\tpur$ & 10~GB & 5m2.111s & \todo{X} & \todo{X} & 0.000s & 0.000s & two pipelines merging to a \tti{comm} \\
sort-sort ~&~ $\tsta, 2\times\tpur$ & 10~GB & 3m8.657s & 154 & \todo{X} & 0.327s & 0.000s & parallelizable \tpur after \tpur \\
shortest-scripts ~&~ $5\times\tsta, 2\times\tpur$ & 85~MB & 3m4.233s & \todo{X} & \todo{X} & 0.000s & 0.000s & long \tsta pipeline ending with \tpur \\
\bottomrule
\end{tabular*}